---
title: "Assignment 5"
author: "Allison Tessman"
date: "2024-02-22"
output:
  html_document: default
  pdf_document: default
---

1. Run the codes in the slides to reproduce the results.

# *Rainfall data: Constant level (no trend) and no seasonality*

```{r}
rainfall <- read.csv("https://bryantstats.github.io/math475/slides/rainfall_london.csv")

# install.packages("forecast")
library(forecast)

#forecasting with single exponential smoothing
ts1 <- ts(rainfall,start=c(1813))
plot(ts1)

#making forecast
ts1_forecast <- HoltWinters(ts1, alpha=TRUE, 
                            beta=FALSE, 
                            gamma=FALSE)
ts1_forecast
plot(ts1_forecast)

#forecasting
ts1_forecast2 <- forecast(ts1_forecast, h=8)
ts1_forecast2
plot(ts1_forecast2)

```

# *Skirts data: Increasing or decreasing trend and no seasonality*

```{r}
skirts <- read.csv("https://bryantstats.github.io/math475/slides/skirts.csv")

#forecasting with single exponential smoothing
ts2 <- ts(skirts,start=c(1866))
plot(ts2)

#making forecast
ts2_forecast <- HoltWinters(ts2, alpha=TRUE, 
                            beta=TRUE, 
                            gamma=FALSE)
ts2_forecast
plot(ts2_forecast)

#forecasting
ts2_forecast2 <- forecast(ts2_forecast, h=19)
plot(ts2_forecast2)

```

# *Souvenir data: Increasing or decreasing trend and seasonality*

```{r}
souvenir <- read.csv("https://bryantstats.github.io/math475/slides/souvenir.csv")

#forecasting with single exponential smoothing
ts3 <- ts(souvenir, frequency=12, start=c(1987,1))
plot(ts3)

#making forecast
ts3_forecasts <- HoltWinters(ts3, alpha=TRUE, 
                            beta=TRUE, 
                            gamma=TRUE)
plot(ts3_forecasts)

#forecasting
ts3_forecasts2 <- forecast(ts3_forecasts, h=48)
plot(ts3_forecasts2)

#log(ts3) example
ts3 <- ts(souvenir, frequency=12, start=c(1987,1))
logts3 <- log(ts3)

ts3_forecasts <- HoltWinters(logts3)
plot(ts3_forecasts)

ts3_forecasts2 <- forecast(ts3_forecasts, h=48)
plot(ts3_forecasts2)

```

2. Find three examples of the below time series and make forecast using Exponential Smoothing

- Time series with no trend and no seasonality:

```{r}
library(readxl)
rainfall_brazil <- read_excel("~/Applied Analystics SAS Prog/Assignments/Assignment 5/rainfall_london.xlsx", 
    sheet = "new_rainfall")

# install.packages("forecast")
library(forecast)

#forecasting with single exponential smoothing
ts1 <- ts(rainfall_brazil,start=c(1813))
plot(ts1)

#making forecast
ts1_forecast <- HoltWinters(ts1, alpha=TRUE, 
                            beta=FALSE, 
                            gamma=FALSE)
ts1_forecast
plot(ts1_forecast)

#forecasting
ts1_forecast2 <- forecast(ts1_forecast, h=8)
ts1_forecast2
plot(ts1_forecast2)

```

- Time series with trend and no seasonality:

```{r}
shorts <- read_excel("~/Applied Analystics SAS Prog/Assignments/Assignment 5/skirts.xlsx", 
    sheet = "new_shorts")

#forecasting with single exponential smoothing
ts2 <- ts(shorts,start=c(1866))
plot(ts2)

#making forecast
ts2_forecast <- HoltWinters(ts2, alpha=TRUE, 
                            beta=TRUE, 
                            gamma=FALSE)
ts2_forecast
plot(ts2_forecast)

#forecasting
ts2_forecast2 <- forecast(ts2_forecast, h=19)
plot(ts2_forecast2)

```

- Time series with trend and seasonality:

```{r}
tickets <- read_excel("~/Applied Analystics SAS Prog/Assignments/Assignment 5/souvenir.xlsx", 
    sheet = "tickets")

#forecasting with single exponential smoothing
ts3 <- ts(tickets, frequency=12, start=c(1992,1))
plot(ts3)

#making forecast
ts3_forecasts <- HoltWinters(ts3, alpha=TRUE, 
                            beta=TRUE, 
                            gamma=TRUE)
plot(ts3_forecasts)

#forecasting
ts3_forecasts2 <- forecast(ts3_forecasts, h=48)
plot(ts3_forecasts2)

#log(ts3) example
ts3 <- ts(souvenir, frequency=12, start=c(1992,1))
logts3 <- log(ts3)

ts3_forecasts <- HoltWinters(logts3)
plot(ts3_forecasts)

ts3_forecasts2 <- forecast(ts3_forecasts, h=48)
plot(ts3_forecasts2)


```

