---
title: "Assignment 8"
author: "Allison Tessman"
date: "2024-03-05"
output:
  html_document: default
  word_document: default
---

# Question 1: Differencing Techniques

- Plot the ACF for the Microsoft stock series for 2023. Investigate the stationarity of the series.

- NOTE: The correlations does not die out to zero, so this series is non-stationary and is not reasonable to model the stock using the AR model.

```{r}
library(quantmod)
library(forecast)
getSymbols("MSFT")

yt = MSFT$MSFT.Open
yt <- yt[index(yt) > "2023-01-01"]

acf(yt, lag.max = 100)

```

- Using the differencing technique to forecast the next value of the stock series.

```{r}
#acf plot for difference series
d_MSFT = ts(as.numeric(diff(yt))[-1])
acf(d_MSFT, lag.max = 100)

ar_MSFT = arima(d_MSFT, order = c(1,0,0))
ar_MSFT

#forecast the next observation
d_n = forecast(ar_MSFT, h = 1)

y_next = d_n$mean + yt[length(yt)]
y_next = as.numeric(y_next)
y_next

# [1] 413.8426
```

# Question 2: AR(p)

- Consider the AMD stock from 2023 to present. Use the PACF plot to suggest an AR(p) model to model the stock. Use the model to make a forecast for the next 5 values. Plot the forecasting.

```{r}
library(quantmod)
library(forecast)
getSymbols("AMD")

yt = AMD$AMD.Open
yt <- yt[index(yt) > "2023-01-01"]
pacf(yt)

# estimate the series using AR(1) model
yt_ar  = arima(yt, order = c(1,0,0))

# plot the estimated series and the original series
yt_predicted <- yt - yt_ar$residuals
plot(yt)
points(yt_predicted, type = "l", 
       col = "red", lty = 2)

yt_forecasts <- forecast(yt_ar, h=5)
plot(yt_forecasts)

```

- Redo the above analysis on another stock.

```{r}
library(quantmod)
library(forecast)
getSymbols("DIS")

yt = DIS$DIS.Open
yt <- yt[index(yt) > "2022-01-01"]
pacf(yt)

# estimate the series using AR(1) model
yt_ar  = arima(yt, order = c(1,0,0))

# plot the estimated series and the original series
yt_predicted <- yt - yt_ar$residuals
plot(yt)
points(yt_predicted, type = "l", 
       col = "red", lty = 2)

yt_forecasts <- forecast(yt_ar, h=5)
plot(yt_forecasts)
```

