---
title: "Assignment 13"
author: "Allison Tessman"
date: "2024-04-09"
output:
  word_document: default
  html_document: default
---

Question 1. Working with a text dataset containing Amazon reviews for an iPhone model. Do the following:

- Create a list of all possible pairs of words (not only adjacent words) that do not contain stopwords and plot the most frequent pairs.

```{r}
library(tidyverse)
library(ggplot2)
library(tidytext)
df <- read_csv("https://bryantstats.github.io/math475/assignments/Amazon_Unlocked_Mobile2.csv")

df %>% 
  group_by(Product_Name) %>% 
  count(sort = TRUE)

df_section_words = df %>% 
  mutate(section = c(1:nrow(df))) %>% 
  unnest_tokens(output = word, input = texts) %>% 
  filter(!word %in% stop_words$word)
           

library(widyr)

# count words co-occuring within sections
word_pairs <- df_section_words %>%
  pairwise_count(word, section, sort = TRUE)

word_pairs

word_pairs %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot() + 
  geom_col(aes(y = fct_rev(fct_reorder(item, -n)), x = n))+
  labs(y = "", x = 'Frequency')
```

- Plot the graph of the pairs of words

```{r}
library(igraph)
pairs_graph <- word_pairs %>%
  filter(n > 50) %>%
  graph_from_data_frame()

library(ggraph)
set.seed(2024)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(pairs_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

```

- Compute the correlations of all the pairs and plot the bar chart of the pairs with the highest correlation

```{r}
word_cors <- df_section_words %>%
  group_by(word) %>%
  filter(n() >= 20) %>%
  pairwise_cor(item = word, feature = section, sort = TRUE)

word_cors

word_cors %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot()+
  geom_col(aes(y = fct_rev(fct_reorder(item, -correlation)), x = correlation))+
  labs(y ='')

```

- Plot the graph of the correlation

```{r}
library(ggraph)
library(igraph)
set.seed(2024)

word_cors %>%
  filter(correlation > .3) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```


Question 2. Do Question 1 on your own text dataset.

```{r}
library(tidyverse)
library(ggplot2)
library(tidytext)

df <- read_csv('~/Applied Analystics SAS Prog/mymath475/CNNtext.csv')

df %>% 
  group_by(id) %>% 
  count(sort = TRUE)

df_section_words = df %>% 
  mutate(section = c(1:nrow(df))) %>% 
  unnest_tokens(output = word, input = highlights) %>% 
  filter(!word %in% stop_words$word)
           

library(widyr)

# count words co-occuring within sections
word_pairs <- df_section_words %>%
  pairwise_count(word, section, sort = TRUE)

word_pairs

word_pairs %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot() + 
  geom_col(aes(y = fct_rev(fct_reorder(item, -n)), x = n))+
  labs(y = "", x = 'Frequency')

library(igraph)
pairs_graph <- word_pairs %>%
  filter(n > 150) %>%
  graph_from_data_frame()

library(ggraph)
set.seed(2024)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(pairs_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

word_cors <- df_section_words %>%
  group_by(word) %>%
  filter(n() >= 20) %>%
  pairwise_cor(item = word, feature = section, sort = TRUE)

word_cors

word_cors %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot()+
  geom_col(aes(y = fct_rev(fct_reorder(item, -correlation)), x = correlation))+
  labs(y ='')

library(ggraph)
library(igraph)
set.seed(2024)

word_cors %>%
  filter(correlation > .75) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```

