---
title: "Assignment 4"
author: "Allison Tessman"
date: "2024-02-15"
output:
  pdf_document: default
  html_document: default
---

# Plot the Airline Passenger data and its STL decomposition

```{r}
data("AirPassengers")
plot(AirPassengers)

ourDecomposition <- stl(AirPassengers, s.window = "periodic")
plot(ourDecomposition)
```

# Plot the Electric Production and its stl decomposition

```{r}
# read data
d <- read.csv('https://bryantstats.github.io/math475/assignments/data/Electric_Production.csv')

# frequency = 4 for quarterly data, 
# frequency = 12 for monthly data, 
# frequency = 1 for yearly data, by default frequency = 1

t1 = ts(d$IPG2211A2N, start = 1985, frequency = 12)

# plot the time series
plot(t1, main = paste0("Electric Production"))

#stl decomposition
ourDecomposition2 <- stl(t1, s.window = "periodic")
plot(ourDecomposition2)
```

# Plot the Electric Production from 1985 to 1995 and its stl decomposition

```{r}
# read data
d <- read.csv('https://bryantstats.github.io/math475/assignments/data/Electric_Production.csv')

# frequency = 4 for quarterly data, 
# frequency = 12 for monthly data, 
# frequency = 1 for yearly data, by default frequency = 1

t1 = ts(d$IPG2211A2N, start = 1985, end = 1995, frequency = 12)

# plot the time series
plot(t1, main = paste0("Electric Production from 1985 to 1995"))

#stl decomposition
ourDecomposition3 <- stl(t1, s.window = "periodic")
plot(ourDecomposition3)
```

# Plot the US Retail Employment and its STL decomposition

```{r}
# read data
d <- read.csv('https://bryantstats.github.io/math475/assignments/data/us_retail_employment.csv')

# frequency = 4 for quarterly data, 
# frequency = 12 for monthly data, 
# frequency = 1 for yearly data, by default frequency = 1

t1 = ts(d$Employed, start = 1990, frequency = 4)

# plot the time series
plot(t1, main = paste0("US Retail Employment"))

#stl decomposition
ourDecomposition4 <- stl(t1, s.window = "periodic")
plot(ourDecomposition4)
```

# Find your own time series that has a seasonal component and plot the classical decomposition (as in Assignment 3) and the STL decomposition.

```{r}
d <- read.csv("~/Applied Analystics SAS Prog/mymath475/daily_csv.csv")

library(zoo)
d$month <- as.yearmon(d$Date)

d <- d[-c(1)]
d$Date <- d$month
d <- d[-c(2)]

#Create time series with data
t1 = ts(d$Date, start = 2000, end = 2010, frequency = 12)
plot(t1, main = paste0("Natural Gas Prices"))

#Additive decomposition
ourDecomposition <- decompose(t1, type="additive")
plot(ourDecomposition)

#Multiplicative decomposition
ourDecomposition <- decompose(t1, type="multiplicative")
plot(ourDecomposition)

#STL decomposition
ourDecomposition <- stl(t1, s.window = "periodic")
plot(ourDecomposition)
```

