---
title: "Final Project"
author: "Allison Tessman"
date: "2024-04-26"
output: html_document
---

Subset the data:

1. Bigrams Analysis (Assignment 12)

Do the following:

- Create a list of pairs of words that may contain stopwords and plot the most frequent pairs.

- Create a list of pairs of words that do not contain stopwords and plot the most frequent pairs.

- Visualize the pairs of words (bigrams)

- Plot the context sentiment analysis

```{r, warning=FALSE}
library(tidyverse)

df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),]  

# only look at sub product with most reviews
library(tidyverse)
library(ggplot2)
library(tidytext)
library(tm)

df %>% 
  group_by(`Sub-issue`) %>% 
  count(sort = TRUE)

# filter out to only include credit reporting (#1 complaints)
df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>% 
  filter(`Sub-issue`=='Information belongs to someone else') %>% 
  rename(texts = `Consumer complaint narrative`)

# create a list of pairs of words
df_bigrams <- df %>%
  unnest_tokens(input = texts, output = bigram, token = "ngrams", n = 2) %>% 
  filter(!is.na(bigram))

df_bigrams %>%
  count(bigram, sort = TRUE)

# top most frequent pairs
df_bigrams %>%
  count(bigram, sort = TRUE) %>% 
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(bigram, n))) +
  geom_col() +
  labs(y = '', x = 'Frequency')

# remove stop words
bigrams_separated <- df_bigrams %>%
    separate(bigram, c("word1", "word2"), sep = " ")

stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xxxx', 'xx'))

bigrams_filtered <- bigrams_separated %>%
    filter(!word1 %in% stop_words$word) %>%
    filter(!word1 %in% stop_words$word) %>%
    filter(!word1 %in% stop_word2$word) %>%
    filter(!word2 %in% stop_word2$word)

# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
    count(word1, word2, sort = TRUE)
bigram_counts

# top most frequent pairs
bigram_counts$pairs = paste0(bigram_counts$word1, " ", bigram_counts$word2)

bigram_counts %>% 
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(pairs, n))) +
  geom_col() +
  labs(y = '', x = 'Frequency')

# visualize the bigrams
library(igraph)
bigram_graph <- bigram_counts %>%
  filter(n > 20) %>%
  graph_from_data_frame()

library(ggraph)
set.seed(2024)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

# create a wordcloud
## create a wordcloud
library(wordcloud) 
bigram_counts %>% 
  with(wordcloud(pairs, n, random.order = FALSE, 
                 max.words = 40, 
                 colors=brewer.pal(8,"Dark2")))

# *********************************************

# Trigrams
df_trigrams <- df %>%
  unnest_tokens(input = texts, output = trigram, token = "ngrams", n = 3) %>% 
  filter(!is.na(word))

df_separated <- df_trigrams %>%
    separate(trigram, c("word1", "word2", "word3"), sep = " ")

stop_word3 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xxxx', 'xx'))

df_filtered <- df_separated %>%
    filter(!word1 %in% stop_words$word) %>%
    filter(!word2 %in% stop_words$word) %>%
    filter(!word3 %in% stop_words$word) %>%
    filter(!word1 %in% stop_word3$word) %>%
    filter(!word2 %in% stop_word3$word) %>% 
    filter(!word3 %in% stop_word3$word)

trigram_count <- df_filtered %>% 
  count(word1, word2, word3, sort = TRUE) %>% 
  unite(word, word1, word2, word3, sep = " ")

## plot the frequency
trigram_count %>% 
  filter(n > 15) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) + 
  geom_col() + 
  xlab(NULL) + 
  coord_flip()

# wordcloud
trigram_count %>% 
  with(wordcloud(word, n, random.order = FALSE, 
                 max.words = 50, colors=brewer.pal(8,"Dark2")))

# ***********************************************

# Content Sentiment Analysis

# words preceded by NOT
bigrams_separated %>%
  filter(word1 == "not") %>%
  count(word1, word2, sort = TRUE)

AFINN <- get_sentiments("afinn")

not_words <- bigrams_separated %>%
  filter(word1 == "not") %>%
  inner_join(AFINN, by = c(word2 = "word")) %>%
  count(word2, value, sort = TRUE)

not_words

# plot sentiment analysis
# NOTE: use of word NOT can cause reversed sentiment and is issue
library(ggplot2)

not_words %>%
  mutate(contribution = n * value) %>%
  arrange(desc(abs(contribution))) %>%
  head(20) %>%
  mutate(word2 = reorder(word2, contribution)) %>%
  ggplot(aes(n * value, word2, fill = n * value > 0)) +
  geom_col(show.legend = FALSE) +
  labs(x = "Sentiment value * number of occurrences. These sentiment should be reversed",
       y = "Words preceded by \"not\"")

# identify words that connect with not, no, never, without
# then reverse the sentiment values of these words
stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx'))
AFINN <- get_sentiments("afinn")

sentiment1 = df %>%
  unnest_tokens(input = texts, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  anti_join(stop_word2) %>% 
  inner_join(get_sentiments("afinn")) %>% 
  filter(!is.na(value)) %>% 
  count(value)

sentiment1 %>% 
  ggplot(aes(x = value, y = n))+
  geom_col()+  
  labs(title = 'Non-context Sentiment Analysis')

negation_words <- c("not", "no", "never", "without")

sentiment2 = bigrams_separated %>%
  filter(word1 %in% negation_words) %>%
  inner_join(AFINN, by = c(word2 = "word")) %>%
  count(word1, word2, value, sort = TRUE) %>% 
  select(value, n) %>% 
  group_by(value) %>% 
  summarise(n = sum(n)) %>% 
  arrange(value)

ggplot()+
  geom_col(data = sentiment1, aes(x =value, y = n))+
  geom_col(data = sentiment2, aes(x =value, y = n), fill = 'red')+
  labs(title = 'Context Sentiment Analysis',
       caption = "Red bars present the reversed sentiment.")
# Red bars present the reversed sentiment!

```


2. Word Pair and Correlation Analysis (Assignment 13)

Do the following:

- Create a list of all possible pairs of words (not only adjacent words) that do not contain stopwords and plot the most frequent pairs.

- Plot the graph of the pairs of words

- Compute the correlations of all the pairs and plot the bar chart of the pairs with the highest correlation

- Plot the graph of the correlation

```{r, warning=FALSE}
library(tidyverse)

df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),]  


# Find item with most reviews
df %>% 
  group_by(`Sub-issue`) %>% 
  count(sort = TRUE)

# filter out to only include credit reporting (#1 complaints)
df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>% 
  filter(`Sub-issue`=='Information belongs to someone else') %>% 
  rename(texts = `Consumer complaint narrative`)

# pairs of words
stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xxxx', 'xx'))

df_section_words = df %>% 
  mutate(section = c(1:nrow(df))) %>% 
  unnest_tokens(output = word, input = texts) %>% 
  filter(!word %in% stop_words$word) %>% 
  filter(!word %in% stop_word2$word)

library(widyr)

# count words co-occuring within sections
word_pairs <- df_section_words %>%
  pairwise_count(word, section, sort = TRUE)

word_pairs

# top most frequent pairs
word_pairs %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot() + 
  geom_col(aes(y = fct_rev(fct_reorder(item, -n)), x = n))+
  labs(y = "", x = 'Frequency')

# graph of word pairs
library(igraph)
pairs_graph <- word_pairs %>%
  filter(n > 40) %>%
  graph_from_data_frame()

library(ggraph)
set.seed(2024)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(pairs_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

# ********************************************

# Pairwise Correlation

# top correlated pairs
word_cors <- df_section_words %>%
  group_by(word) %>%
  filter(n() >= 20) %>%
  pairwise_cor(item = word, feature = section, sort = TRUE)

word_cors

word_cors %>% 
  head(20) %>% 
  mutate(item = paste0(item1, " ", item2)) %>% 
  ggplot()+
  geom_col(aes(y = fct_rev(fct_reorder(item, -correlation)), x = correlation))+
  labs(y ='')

# graph of the correlations
library(ggraph)
library(igraph)
set.seed(2024)

word_cors %>%
  filter(correlation > .65) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()

```


3. tf-idf plots (Assignment 14)

Do the following:

- Define your own documents for the analysis.

- Plot the term frequency in the documents

- Plot the tf-idf in the documents.

```{r, warning=FALSE}
library(tidyverse)

df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),] 

# define documents
df %>% 
  group_by(`Sub-issue`) %>% 
  count(sort = TRUE)

# filter out data set for 2 items to analyze
df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>% 
  filter(`Sub-issue` %in% c('Information belongs to someone else', 'Reporting company used your report improperly')) %>% 
  rename(texts = `Consumer complaint narrative`, 
         document = `Sub-issue`)

# plot the term freqency for each document
stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx'))

df_words = df %>% 
  unnest_tokens(input = texts, output = word) %>% 
  count(document, word, sort = TRUE) %>% 
  anti_join(get_stopwords()) %>% 
  anti_join(stop_word2)

total_words <- df_words %>% 
  group_by(document) %>% 
  summarize(total = sum(n))

df_words <- left_join(df_words, total_words)

df_tf = df_words %>% 
  group_by(document) %>% 
  mutate(tf = n/total)

df_tf %>% 
  group_by(document) %>% 
  slice_max(tf, n = 5) %>% 
  ungroup() %>%
  ggplot(aes(tf, fct_reorder(word, tf), fill = document)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~document, ncol = 2, scales = "free") +
  labs(x = "tf", y = NULL)

# plot the tf-idf for each doc
df_tf_idf <- df_words %>%
  bind_tf_idf(word, document, n)

df_tf_idf

library(forcats)

df_tf_idf %>%
  group_by(document) %>%
  slice_max(tf_idf, n = 15) %>%
  ungroup() %>%
  ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = document)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~document, ncol = 2, scales = "free") +
  labs(x = "tf-idf", y = NULL)

```


4. Text Clustering (Assignment 15)

Do the following:

- Print out a sample of the Document Term Matrix using term frequency

- Print out a sample of the Document Term Matrix using tf-idf

- Cluster the words in the dataset using k-means.

- Plot the words frequency of the words in each clusters

- Plot the frequency of the part of speech of all the words in the dataset.

```{r, warning=FALSE}
# document term matrix- term frequency
library(tidytext)
library(tm)
library(wordcloud)
library(tidyverse)

df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),] 

df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>%
  rename(texts = `Consumer complaint narrative`, 
         document = `Sub-issue`)

# create the DTM
df_tm <- df %>% 
  unnest_tokens(output = word, input = texts) %>% 
  anti_join(get_stopwords()) %>% 
  anti_join(tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx',  as.character(c(1:100)))))

word_freq <- df_tm %>% 
  group_by(document) %>% count(word, sort = TRUE)

df_dtm <- word_freq %>% 
  cast_dtm(document = document, term = word, n)

tm::inspect(df_dtm)

# DTM- tf-idf
df_dtm <- weightTfIdf(df_dtm)

tm::inspect(df_dtm)

# *****************************************

# Clustering

# transpose data so terms are rows of data
# df_dtm <- removeSparseTerms(df_dtm, 0.5)
kmeans.data <- as.matrix(t(df_dtm))

kfit <- kmeans(kmeans.data, 3)

# plot word cloud
df %>%
  unnest_tokens(input = texts, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  anti_join(tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx',  as.character(c(1:100))))) %>% 
  count(word, sort = TRUE) %>% 
  with(wordcloud(word, n, random.order = FALSE, 
                 max.words = 50, colors=brewer.pal(8,"Dark2")))

# plot cluster 1
k = 1

cluster = names(kfit$cluster)[kfit$cluster==k]
cluster = as_tibble(cluster) %>% 
  rename(word = value)

cluster = word_freq <- df_tm %>% 
  count(word, sort = TRUE) %>% 
  inner_join(cluster, by = 'word')

library(wordcloud)
set.seed(2024)
cluster %>% with(wordcloud(word, n, max.words = 50, random.order = FALSE, rot.per = 0.35, 
    colors = brewer.pal(8, "Dark2")))

# plot cluster 2
k = 2

cluster = names(kfit$cluster)[kfit$cluster==k]
cluster = as_tibble(cluster) %>% 
  rename(word = value)

cluster = word_freq <- df_tm %>% 
  count(word, sort = TRUE) %>% 
  inner_join(cluster, by = 'word')

library(wordcloud)
set.seed(2024)
cluster %>% with(wordcloud(word, n, max.words = 50, random.order = FALSE, rot.per = 0.35, 
    colors = brewer.pal(8, "Dark2")))

# plot cluster 3
k = 3

cluster = names(kfit$cluster)[kfit$cluster==k]
cluster = as_tibble(cluster) %>% 
  rename(word = value)

cluster = word_freq <- df_tm %>% 
  count(word, sort = TRUE) %>% 
  inner_join(cluster, by = 'word')

library(wordcloud)
set.seed(2024)
cluster %>% with(wordcloud(word, n, max.words = 50, random.order = FALSE, rot.per = 0.35, 
    colors = brewer.pal(8, "Dark2")))

# *******************************************

# Part of Speech

# map words into its part of speech (noun, ver, etc.)
library(tidyverse)
library(tidytext)
library(tm)
library(wordcloud)

df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),] 

df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>%
  rename(texts = `Consumer complaint narrative`, 
         document = `Sub-issue`)

stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx'))

df %>% 
    unnest_tokens(input = texts, output = word) %>% 
    count(word, sort = TRUE) %>% 
    anti_join(get_stopwords()) %>% 
    anti_join(stop_word2) %>% 
    inner_join(parts_of_speech) %>%                   # join POS
    count(pos) %>%                                    # count
    mutate(prop=n/sum(n)) %>% 
    slice_max(prop, n = 15) %>%
    ggplot()+geom_col(aes(y = pos, x = prop), position = 'dodge')

# doc based parts of speech
df <- read_csv("~/Applied Analystics SAS Prog/mymath475/final_complaints.csv")

# subset the first 500 rows of the data
df = df[c(1:500),]  

# OR

# a random sample of 500 rows from the data
#df = df[sample(nrow(df), 500),] 

df = df %>% 
  select(`Sub-issue`, `Consumer complaint narrative`) %>%
  rename(texts = `Consumer complaint narrative`, 
         document = `Sub-issue`)

df <- df %>% 
  filter(grepl("Information belongs to someone else|Reporting company used your report improperly", document))

stop_word2 = tibble(word = c(letters, LETTERS, "oh", 'just', 'xx', 'xxxx'))

df %>% 
    unnest_tokens(input = texts, output = word) %>% 
    group_by(document) %>% 
    count(document, word, sort = TRUE) %>% 
    anti_join(get_stopwords()) %>% 
    anti_join(stop_word2) %>% 
    inner_join(parts_of_speech) %>%                   # join POS
    count(pos) %>%                                    # count
    mutate(prop=n/sum(n)) %>% 
    slice_max(prop, n = 15) %>%
    ggplot()+geom_col(aes(y = pos, x = prop, fill =  document), position = 'dodge')

```


